### 2021-10-11 plos_latex_template.tex
AST: i was checking the output of the PDF and deleted extra tex files that overleaf saves separately and @rociojoo had added (sorry!)  nothing changes technically. the output seems to be off when you look at it on overleaf but if you open the link to the actual PDF it's ok. you can add them againif you want, Rocío but overleaf does not do anything to them.

In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for global conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. 
% Rocío: we can add our Nature Correspondence letter to the citations later on. 
Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020}.
% Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location. 
% Rocío: main idea of the paragraph: online conferences are more inclusive
% Liz: but only if you make them that way. The R conference was a standout and innovative compared to others I've considred attending since Covid.

Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}. 
However, social events in in-person conferences can be exclusive (e.g. galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. % Rocío: any reference to back the last statement? The CoC guide? ast: yes
Organizers should invest time in creating opportunities to meet and bond virtually. 
An online event is a great opportunity to create spaces for people who do not enjoy in person networking. 
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people's limits, preferences, and remember that `the usual' does not necessarily work with everyone in your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.
Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
% Rocío: Networking paragraph.

@@

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
Needless to say, an inclusive online component of a hybrid conference is not a free pass to have an exclusionary in-person component. % Rocío: I know this is a bit aggressive so I wouldn't mind if anybody helps with the wording. 
% Liz: I like it, but you could say
% an inclusive online component doesn't eliminate the need to organize an inclusive in-person component.

For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Hybrid conferences paragraph.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
Importantly, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general. 
% Liz: I don't think it's actually ironic. maybe surprising or counterintuitive? ast: absolutely! i put importantly at the end to highlight that without opposition
If the conference is in person, the venue must be accessible for people with mobility limitations. 
% Liz: let's stay away from deficit-model language like "limitations." Possible aternatives:
% people who use wheelchairs or other assistive mobility devices
% ... must be physically accessible for disabled people
Accommodations such as a quiet space for neurodivergent people should also be provided.
% Liz: I hate "accommodateions because it sounds like something optional or for comfort or luxurious Consider something like "facilities"?

Presenters should always speak into a microphone to make it easier for people who are hard of hearing and for captioners or interpreters to listen to them. 
% Liz: avoid using disabilities as nouns like -- "the blind", "the deaf" etc. I woul duse "people who are deaf or hard-of-hearing" above. There's controversy about whether the word person
% should come first ("person who is blind") or the identity should come first ("blind person:) but the person should always be acknowledged (like "participants with disabilities", blind scientist", etc.. It's kind of objectifying without a person-related word.
Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. 
% Liz: not just alt-text, the whole site should be screen-reader accessible so:
% communication about the conference should be fully screen-reader accessible including alt-text, and any videos should have both captions and sound.

Platforms for conference registration and abstract submission, websites, and chat platforms--if used--should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning for presentations should be available in more than one language if possible.
Do not forget about making your social events accessible (e.g. art shows and movies) and making sure that the materials presented are accessible for all. 
These aspects should be tested well in advance of going live.

Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.  
Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).

% Liz: It is essential to provide an accessible version of the slides for people who won't be seeing them One of the most important assumptions about presentations are that they will be both seen and heard. Usually, neither of these channels is a complete source of information about the presentation. For example, when discussing computer programming, code is usually displayed on slides while the presenter talks about them. They seldom read the details of what appears on the slides in enough detail for a blind person to learn, and a deaf person will also miss some explanation. Presenters  should provide as much information as possible on both channels, but in order to compensate for the fact that this is not common practice, blind people need to be able to read an accessible version of the slides at the same time they listen to the presentation. It is critical to make these materials available before the conference starts..

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs in an in-person conference.

On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g. choosing between captions and interpreters). 
Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
% Liz: maybe strengthen the last phrase to use their expertise to solve accessibility problems and collaborate on decisions about accessibility practices.

% Liz: As design justice author Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".  %Rocío: maybe slightly rephrase the last sentence since it seems to imply that everybody knows Sasha Costanza-Chock.
% not sure if that's a good characterization-- I am just learning about her right now.

@@

Meeting the need 
% Liz: People with disabilities don't want to be considered "needy". We reject the term "special needs" in part because it frames the person with the disability as being the burden causing the accommodation, where maybe it's the responsibility of society to be open to everyone. So I wonder if it's the same for people who speak other languages. Maybe they are not needy or inferior.  Or maybe it's wrong for me to apply disability rights philosophy to language.
% The linguistic diversity of conference participants is often overlooked,
% resulting in missed opportunities for them, as well as for the conference
% to benefit from their contributions.
of linguistically diverse audience is an important aspect in inclusion that is often overlooked. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

@@

% Rocío: Pasting Batool's text in her 10 simple rules because there are very valuable things we could use: "Rule: Meeting the need of linguistically diverse audience. Most international conferences are held in English. However, English is only spoken by 4.83 percent of the world population. If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.    If possible, ensure that some sessions are delivered by languages other than English"

@@

As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct and accessibility guidelines, and providing fee waivers and financial support. People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory. If possible, include representatives from these communities within the organizing team (\textbf{Rule \ref{rule_inclusion}}). Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.

Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}). Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusions. Be careful with the language you use in promoting to the conference and how you talk about minoritized people, assertively set tone of inclusion. In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept feedback without being offended.

As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

% Yani: We need to work a little more here, perhaps with some examples.
% Rocío: could add some ideas from here 
% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
% Batool: I will work on this rule to include more details by 25th of Sept

%sara: this fits better in the communication rule
%adding here from Liz's rule but maybe this could go in communication in general
%In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept feedback without being offended.

@@

%Inclusion comes with a cost

Conference budgets are limited and rely mostly on sponsorship and attendance fees. Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference. It is important to estimate the costs for these inclusive practices and define your priorities in advance. This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership training, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}). It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team. Don't be afraid to innovate and resist the institutional inertia. "This has always been like that" will not take the conference towards structural change.
%should we talk about the pandemic? something like: 
With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.

%Set fair registration rates
Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee. The definition of such positions and the associated work conditions vary for each country. Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere. In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies. You can adjust the registration fee by the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example). In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach. Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.

Offering scholarships or grants to attend the conference is a common way to boost participation from different groups, and some conferences ask for cover letters or applications to them. These applications may be counterproductive, if they attract people that do not need the money but apply to enrich their CVs, and can filter out people who do not feel entitled to earn them.  
Applying for loans, grants, and scholarships may be an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way-- when someone says they need the support, it is best to give them the waiver rather than second guess their eligibility. If the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.

@@

The ten rules stated here can be adapted depending on the conference format and settings.
When organizing useR!2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
We had to let go of certain aspects that were important to many of us. 
From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
Also, keep working to improve inclusion in the next edition of the conference! to document and balance your work, design surveys, ask for feedback from attendees and organizers. Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team. Always remember inclusion is an ambition that organizers can and should continually pursue by allowing room for learning and improvement.
If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
And you will make a change towards healthier, stronger, and more inclusive communities.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR!2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!2021 and supporting us through the process.

>>>>

In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. 
% Rocío: we can add our Nature Correspondence letter to the citations later on. 
Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020}.
% Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location. 
% Rocío: main idea of the paragraph: online conferences are more inclusive
% Liz: but only if you make them that way. The R conference was a standout and innovative compared to others I've considred attending since Covid.

Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}. 
However, social events in in-person conferences can be exclusive (e.g. galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. Organizers should invest time in creating opportunities to meet and bond virtually. 
An online event is a great opportunity to create spaces for people who do not enjoy in-person networking. 
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people's limits, preferences, and remember that 'the usual' does not necessarily work for everyone in your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.
Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
% Rocío: Networking paragraph.

@@

Alternatively, a conference could have a hybrid format with an in-person and an online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%ast: I added Liz's suggestion and rephrased slightly to convey that online is not automatically inclusive
For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Hybrid conferences paragraph.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when the organizing team implements other inclusive practices, [ö improve this] the concern about the participation of people with disabilities tends to be partial, and is one of the lesser known attended features of inclusiveness.

Importantly, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

If the conference is in person, the venue must be accessible for people who use wheelchairs, have important signs in Braille, and a sound system compatible with hearing devices, just to name a few important features. Presenters should always speak into a microphone. Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options. 
%ast: I removed the captioners part because it goes slightly off-topic and the sound system is apart for them when in person. 
Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website. Likewise, any videos should have both captions and sound.

Platforms for conference registration and abstract submission, websites, and chat platforms --if used-- should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning or interpretation for presentations should be available in more than one language if possible. Sign languages are more complicated, because the vary from region to region, but if your conference has a clearly regional scope, a specific sign language (such as ASL, American Sign Language, or LIBRAS, the Brazilian Sign Language) can be adopted as official.

Do not forget about making your social events accessible and making sure that the materials presented are accessible for all. This includes physical activities that should not restrict participation based on body type or ability.

These aspects should be tested well in advance of going live.

%teach about accessibility
Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have. Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, in screen-reader-frendly formats (e.g., R Markdown, HTML, or \TeX{} files). Accessibility in slides includes alternative text, and written code instead of print screen.

 If presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}, (see Rule \ref{rule_financial} about allocating resources). Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.

On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g. choosing between captions and interpreters). 
Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot accessibility problems and solve them, and collaborate on decisions about accessibility practices (see \textbf{Rule \ref{rule_organizing_team}}). Simulating disability is not a substitute for actual work with the people directly affected by the decision outcomes\cite{costanzachockDesign2020}

@@

% Liz: People with disabilities don't want to be considered "needy". We reject the term "special needs" in part because it frames the person with the disability as being the burden causing the accommodation, where maybe it's the responsibility of society to be open to everyone. So I wonder if it's the same for people who speak other languages. Maybe they are not needy or inferior.  Or maybe it's wrong for me to apply disability rights philosophy to language.
The linguistic diversity of conference participants is often overlooked, resulting in missed opportunities for them, as well as for the conference to benefit from their contributions. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

@@

%ast: I think ome ideas of the first and second paragraphs can be joined because some parts are in both (intimidation, the importance of not missing opportunities) and maybe some practical tips can be made? Do we add an invitation to "bear" with subtitles and other languages to the English monolingual audience? an invitation to be respectful of "bad English", from submissions to presentations? 
%in person: live interpretation is doable. online: translated captions, pre-recorded talks, english trancripts

@@

As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines [*here] and advertising support for attendees, such as fee waivers and financial support. People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory. If possible, include representatives from these communities within the organizing team (\textbf{Rule \ref{rule_inclusion}}). %ast: i think this sentence can go out, plus these communities 
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.

Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}). Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.%ast this explicit description seems to belong to the previous paragraph, maybe where I marked [*here]?
Be careful with the language you use in promoting the conference and how you talk about minoritized people, assertively set tone of inclusion. In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept feedback without being offended.

As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)

% Yani: We need to work a little more here, perhaps with some examples.
% Rocío: could add some ideas from here 
% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
% Batool: I will work on this rule to include more details by 25th of Sept 
% ast: Thanks Batool! i added a couple of comments.

@@

Conference budgets are limited and rely mostly on sponsorship and attendance fees. Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference. It is important to estimate the costs for these inclusive practices and define your priorities in advance. This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}). It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team. Don't be afraid to innovate and resist the institutional inertia. "We have never payed for this, and this has always been like that" will not take the conference towards structural change.
%ast should we talk about the pandemic? something like: 
With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.

Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee. The definition of such positions and the associated work conditions vary for each country. Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere. In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies. You can adjust the registration fee by the cost of living in each country using conversion factors (e.g., from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example). In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach. Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.

Offering scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and some conferences ask for cover letters or applications to them. These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. For online conferences, granting fee waivers is easier, and diversity scholarships can be replaced with broader participation support.
However, these applications may be counterproductive if they attract people that do not need the money but apply to enrich their CVs, and filter out people who do not feel entitled to earn them. Applying for loans, grants, and scholarships is an emotionally demanding task. Simplify applications as much as you can and be specific about the target audience of these programs. The broader the support, the simpler the application.

@@

The ten rules stated here can be adapted depending on the conference format and settings.
When organizing useR! 2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
We had to let go of certain aspects that were important to many of us. 
From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
Also, keep working to improve inclusion in the next edition of the conference! to document and balance your work, design surveys, ask for feedback from attendees and organizers. Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team. Always remember inclusion is an ambition that organizers can and should continually pursue by allowing room for learning and improvement.
If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
And you will make a change towards healthier, stronger, and more inclusive communities.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR! 2021 and supporting us through the process.
